\section{The first cause and the main theorem}\label{sec:firstcause}

Every reduction of \cref{sec:reduction} presses inward, toward a single open node
$\code{TheLastStep00Obligation}$, narrowed moreover to the scale $A \ge 5$. There is nowhere
further inward to go. The present section turns around and looks outward, to answer three
questions. Why is it legitimate to accept the last node as an external first cause rather than as a
hidden theorem? What does such an architecture prove \emph{unconditionally}, with no axiom at all?
And, deepest, \emph{why} can the node not be closed from inside --- why is knowledge of its cause
unattainable for an observer living inside the system? The answer to the third question is our main
theorem.

\subsection{The cause cannot be derived from inside}

With an open node one may proceed in two ways: leave it open (honest, but the story stops), or
accept it (and everything downstream closes). The second path is legitimate under exactly one
condition --- the acceptance must be \emph{visible}: one may declare an explicit assumption and
trace what rests on it, but one must never pass the unproven off as proven. It turns out there is no
freedom in \emph{how} to accept. To accept the node from inside --- to exhibit its internal
self-grounding --- is impossible, and this is a theorem.

Call an internal proof that crosses its own boundary in order to ground that boundary an
\code{InternalSelfDerivationOfStep00CausalClosure}. Such a construction builds a forbidden concrete
Euclidean engine, $\code{InternalSelfDerivationOfStep00CausalClosure} \to
\code{SomeConcreteEuclideanEngine}$; since there are no engines (the acyclicity of the rank
$\code{lexRank}$, \cref{sec:engine}), it cannot exist.

\begin{theorem}[\code{no\_internalSelfDerivation\_step00CausalClosure}]\label{thm:no_internalSelfDerivation_step00CausalClosure}
\green{} The causal boundary does not self-certify:
$\lnot\,\code{InternalSelfDerivationOfStep00CausalClosure}$.
\end{theorem}

\emph{Proof idea.} The self-certification is identified, step by step under machine check, with the
construction of an infinitely descending clean chain --- exactly a Euclidean engine. No engine
exists, so no self-certification does either.

The same statement in the language of a beginning: a self-caused first frame, an internalised event
$0 \to 1$, cannot exist in a stable engine-free architecture.

\begin{theorem}[\code{no\_internalisedHorizonBoundary}]\label{thm:no_internalisedHorizonBoundary}
\green{} An internalised causal horizon boundary does not exist:
$\lnot\,\code{InternalisedStep00HorizonBoundary}$.
\end{theorem}

\begin{remark}
By itself the prohibition of self-certification is close to tautological: proving $P$ while crossing
the boundary of $P$ collapses into $P \wedge \lnot P$. What is substantive is the
\emph{identification of form} --- that the attempt at self-grounding is already a burnt engine
construction, not some new impossibility. Physically it is the perpetual engine of the first kind:
self-grounding would extract the work of derivation from nothing. The rank $\code{lexRank}$ is a
conserved quantity with a strict floor, and the loop breaks against that floor. Hence the foundation
must be brought in from outside.
\end{remark}

\subsection{An axiom accepted intentionally: the event \texorpdfstring{$0\to1$}{0->1}}

Since the cause cannot be laid from inside, we lay it from outside, deliberately, as the single
axiom of the entire development, \axiom{}. Its type $\code{Step00FirstCause}$ is the structure of the
event $0 \to 1$, the transition from non-being to the first causal frame. Two of its five fields are
pure markers carrying only $\mathrm{True}$ --- $\code{origin}$ (the singularity $0$, where there is
no internal language yet) and $\code{firstFrame}$ (the first frame $1$). The remaining three are the
substantive causal boundaries, one per branch of the programme:

\begin{definition}[the three causal boundaries of \axiom{}]\label{def:step00FirstCause}
$\code{Step00FirstCause}$ carries $\code{causalBoundary}$, the twin boundary --- exactly the open
node $\code{TheStrictLastStep00Obligation}$; $\code{riemannBoundary}$, the Riemann law of
manifestation of a zero (\cref{sec:riemann}); and $\code{nsBoundary}$, the Navier--Stokes gate law of
the energy balance (\cref{sec:ns}). All mathematical weight lives in these three fields; the markers
assert nothing.
\end{definition}

The framing changes the provenance of the result, not its strength: the decree is exactly the sum of
what is put into it.

\begin{theorem}[\code{step00FirstCause\_iff\_causalClosure}]\label{thm:step00FirstCause_iff_causalClosure}
\green{} The first cause is equivalent to the conjunction of its boundaries: $\code{Step00FirstCause}
\leftrightarrow (\text{node} \wedge \text{Riemann law} \wedge \text{NS gate})$.
\end{theorem}

The axiom is locked in a \emph{quarantine module}, and the quarantine is a machine procedure: a
separate verifier marks every declaration depending on \axiom{} as \code{AXIOM-TAINTED}, and none has
leaked into the green line. The former closure axiom is now a theorem obtained by projection,
$\code{step00CausalClosure} := \axiom{}.\code{causalBoundary}$. Along the twin branch it yields the
infinitude of the lower twin centres --- \emph{conditionally}.

\begin{theorem}[\code{twinLowersInfinite\_from\_step00CausalClosure}]\label{thm:twinLowersInfinite_from_step00CausalClosure}
\yellow{} Under the accepted boundary, $\code{TwinLowers.Infinite}$.
\end{theorem}

\emph{This is a conditional derivation, not a proof of the twin-prime conjecture.} The decree is
moreover no weaker than its consequence: it exhibits a twin centre above any threshold,
$\forall M_0,\ \exists m,\ M_0 < m \wedge \code{TwinCenterZ}\,m$
(\code{causalClosureAxiom\_asserts\_twins\_at\_every\_scale}, \yellow{}). Accepting the axiom
\emph{is} accepting the twins; it creates no free strength. The axiom-free remainder is honest to
the point of tautology.

\begin{theorem}[\code{nonAxiomaticRemainingObligation\_iff\_lastStep00Obligation}]\label{thm:nonAxiomaticRemainingObligation_iff_lastStep00Obligation}
\green{} The axiom-free remaining obligation equals the old node:
$\code{Step00NonAxiomaticRemainingObligation} \leftrightarrow \code{TheLastStep00Obligation}$
\red{}. No wrapper lowers the open content below the node; $\code{twin\_prime\_conjecture}$ remains a
$\code{sorry}$ and does not close through the quarantine.
\end{theorem}

\subsection{Epistemics: the cause exists, cannot be known, and knowledge breaks everything}

Write $\code{InternalKnowledgeOfCause}$ for the internal derivation of the cause. Three things can be
said. It exists --- that is \axiom{}, \yellow{}, accepted intentionally. It cannot be known --- and
this is a theorem, \green{}, not a caveat.

\begin{theorem}[\code{cause\_unknowable}]\label{thm:cause_unknowable}
\green{} Internal knowledge of the cause is impossible:
$\lnot\,\code{InternalKnowledgeOfCause}$.
\end{theorem}

\emph{Proof idea.} To know the cause from inside would be to derive it from inside, which builds a
perpetual engine ($\code{knowledge\_builds\_perpetualEngine}$:
$\code{InternalKnowledgeOfCause} \to \code{SomeConcreteEuclideanEngine}$); no engine exists. The
unknowability of the cause is not a weakness but a conservation law.

Knowledge, were it possible, would be explosive: it entails $\lnot\,\code{TwinLowers.Infinite}$
($\code{knowledge\_finitizes\_twins}$) \emph{and} $\code{TwinLowers.Infinite}$
($\code{knowledge\_proves\_anything}$), both \green{}, both \emph{ex falso} through the impossible
engine. The substantive, non-explosive residue is the dichotomy
$\lnot\,\code{InternalKnowledgeOfCause} \vee \lnot\,\code{TwinLowers.Infinite}$
($\code{unknowable\_or\_twins\_finite}$, \green{}), whose left disjunct is
\cref{thm:cause_unknowable}: we are always in the left world.

The load-bearing content is not the explosion but an axiom-free lemma --- the heart of the causal
line --- in which the hypothesis ``no engines'' is consumed substantively rather than vacuously.

\begin{theorem}[\code{twins\_infinite\_of\_noEngine\_and\_boundary}]\label{thm:twins_infinite_of_noEngine_and_boundary}
\green{} The absence of engines together with the causal boundary entails the infinitude of the
twins:
\begin{equation}\label{eq:twins_infinite_of_noEngine}
\lnot\,\code{SomeConcreteEuclideanEngine} \to \code{Step00CausalClosureAxiom} \to
\code{TwinLowers.Infinite}.
\end{equation}
\end{theorem}

\begin{proof}
Suppose the twins are finite; then a last boundary $M_0$ has none above it. From this finite
boundary one builds an infinite family of generated flows; the finite key is forced by pigeonhole to
collide; from the collision a concrete witness engine is assembled. That witness is exactly what the
hypothesis $\lnot\,\code{SomeConcreteEuclideanEngine}$ kills --- not an empty explosion but a direct
hit. The hypothesis is consumed substantively.
\end{proof}

Instantiating \cref{eq:twins_infinite_of_noEngine} by the $\code{lexRank}$-supplied absence of
engines and by the boundary from the axiom gives the twins by decree.

\begin{theorem}[\code{twins\_because\_unknowable}]\label{thm:twins_because_unknowable}
\yellow{} $\code{TwinLowers.Infinite}$, with the derivation visibly passing through the same ``no
engines'' that gives unknowability.
\end{theorem}

\begin{remark}
``The twins are infinite because the cause cannot be known'' means \emph{through a common cause and
a load-bearing lemma}, not ``unknowability alone proves infinitude''. All the weight of
\emph{existence} still rests on the accepted boundary: unknowability without the boundary yields no
twins, or they would already be proven.
\end{remark}

\subsection{The two walls}

There is a second, independent reason the twins escape the observer --- cognitive rather than
causal, and unconditional (its core relies only on $\code{propext}$). Fix a correct certificate
$\code{FiniteSystemKnowsTwin}$ depending only on the finite view of level $A$. Then finite knowledge
is knowledge about a class: a finite system knows a twin $B$ only if its entire sieve-equivalence
class consists of twins ($\code{knowledge\_forces\_pure\_class}$, \green{}); a twin in a mixed class
is invisible in principle ($\code{mixed\_class\_twin\_unknowable}$, \green{}); and if the classes are
tail-mixed, infinitude cannot be certified at all
($\code{infinitude\_unknowable\_of\_eventually\_mixed}$, \green{}), with the trivial view giving the
sharpest unconditional case ($\code{trivialView\_infinitude\_unknowable}$, \green{}). That the
classes of a \emph{non-trivial} view actually mix at concrete levels is an arithmetic input \red{};
the structural theorems and the myopic instantiation are unconditional. The two walls close into one
statement.

\begin{theorem}[\code{two\_walls\_one\_nature}]\label{thm:two_walls_one_nature}
\green{} From inside a finite view a twin with a mixed class cannot be seen, and from inside the
system the first cause cannot be seen:
\[
\bigl(\forall S,A,\mathrm{Cert},B,\mathrm{bad}:\ \code{SieveEquivalent}\,S\,A\,B\,\mathrm{bad} \to
\lnot\,\code{IsTwin}\,\mathrm{bad} \to
\lnot\,\code{FiniteSystemKnowsTwin}\,S\,A\,\mathrm{Cert}\,B\bigr)
\wedge \lnot\,\code{InternalKnowledgeOfCause}.
\]
\end{theorem}

Both walls are one thing: a finite system cannot cross its own boundary. The finite view of level
$A$ is a cognitive horizon, just as the strict order of traversal is a causal one.

\subsection{The main theorem: higher-energy incompatibility}

Everything gathers into a single statement that glues the causal wall to the cognitive wall,
entirely in the green core.

\begin{theorem}[\code{higherEnergyIncompatibility\_main}]\label{thm:higherEnergyIncompatibility_main}
\green{} (no axioms) Five faces of one incompatibility hold conjointly:
(1)~knowledge of the first cause from inside builds a perpetual engine;
(2)~therefore the first cause is unknowable from inside;
(3)~a finite view knows a twin only if its entire class consists of twins;
(4)~under tail mixing of the classes, infinitude cannot be certified from inside;
(5)~the load-bearing face --- this very incompatibility (no engines $=$ the cause cannot be known),
together with the accepted boundary, entails $\code{TwinLowers.Infinite}$.
\end{theorem}

Faces (1)--(4) are the two walls side by side; face (5) is the bridge from them to the conclusion.
Read energetically: knowing from inside costs a perpetual engine, which does not exist, while the
infinitude of the twins is external knowledge, paid for by the first cause. Instantiating face (5) by
decree gives the corollary --- yellow, above a green core.

\begin{corollary}[\code{higherEnergyIncompatibility\_twins}]\label{thm:higherEnergyIncompatibility_twins}
\yellow{} $\code{TwinLowers.Infinite}$. \emph{This is not a proof of the twin-prime conjecture};
$\code{twin\_prime\_conjecture}$ remains a $\code{sorry}$ that does not close through the quarantine.
\end{corollary}

\begin{remark}
The name is not decoration. Landauer's principle makes information physical --- erasing a bit costs
at least $kT\ln 2$, and Maxwell's demon respects the second law precisely because its knowledge must
be paid for. Our theorem is a discrete sibling: knowing the cause from inside would cost infinite
work (a perpetual engine), hence is forbidden by the same conservation law; the infinitude of the
twins is not free, its price being knowledge carried outside, that is, an axiom. This gives the three
worlds the decree draws out. If the node is true, the axiom is superfluous and the twins are in fact
proven; if the node is independent, the theory is consistent but this cannot be known from inside; if
the node is refutable, the quarantine is inconsistent and a tripwire
($\code{quarantine\_inconsistent\_if\_node\_refuted}$, \yellow{}) fires at once, devaluing every
tainted declaration. Both internal resolutions cost a perpetual engine, which is why the boundary can
be accepted only from outside.
\end{remark}
