\section{Yang--Mills and Hodge}\label{sec:ymhodge}

Two Clay problems~\cite{clay-ym,clay-hodge} read, in the language of this article, as
prohibitions of a perpetual engine. In Yang--Mills the forbidden object is a massless spectrum:
an infinite tower of ever-cheaper excitations above the vacuum. In Hodge it is an unpaid charge
descending forever along a natural height. In both branches we prove the structural half that does
not depend on the concrete analytic object mathlib does not have, and we name the red data-anchor
that alone remains. The Lean sources are \srcpath{Engine/YangMillsFront.lean} and
\srcpath{Engine/HodgeFront.lean}.

\subsection{Yang--Mills: masslessness as a perpetual engine}

The mass-gap problem asks us to prove that a quantum gauge theory has a \emph{gap}: a minimal
positive price of excitation above the vacuum. Its negation is \emph{masslessness}, and the reading
of the branch is that masslessness is a perpetual engine --- work drawn from the vacuum with no
lower limit. We are emphatic: we do \emph{not} solve the Clay problem and do not declare it solved.
There is no gauge theory in mathlib --- no Hamiltonian, no genuine spectrum. We work with an
abstract \emph{spectral model} \code{SpectralModel} (a set of energy levels
$\mathrm{Energy}\subseteq\R$ with vacuum $0$ and nonnegativity) and prove the theory-independent
half.

On it, the \emph{mass gap} $\mathrm{MassGap}(S)$ and \emph{masslessness} $\mathrm{Gapless}(S)$ are
exact negations of one another.

\begin{theorem}[\code{not\_massGap\_iff\_gapless}]\label{thm:not_massGap_iff_gapless}
\green{}\ The absence of a gap is equivalent to masslessness, and without any use of choice:
\[
  \neg\,\mathrm{MassGap}(S)\iff\mathrm{Gapless}(S),
\]
where $\mathrm{MassGap}(S):=\exists\,\Delta>0,\ \forall E\in S.\mathrm{Energy},\ E\neq 0\Rightarrow\Delta\le E$
and $\mathrm{Gapless}(S):=\forall\varepsilon>0,\ \exists E\in S.\mathrm{Energy},\ 0<E<\varepsilon$.
\end{theorem}

What makes masslessness an engine is that it builds a \emph{ladder} --- a halving chain of positive
levels, dropping by at least half at each step, $2\cdot E(n+1)\le E(n)$. This is exactly the
$\R$-counterexample of positive work with no lower limit, the infinite multiplicative descent in
the reals against which acyclicity is powerless.

\begin{theorem}[\code{gapless\_iff\_nonempty\_ladder}]\label{thm:gapless_iff_nonempty_ladder}
\green{}\ Masslessness is equivalent to the existence of a ladder: $\mathrm{Gapless}(S)$ holds
if and only if the type $\mathrm{GaplessLadder}(S)$ of ladders is nonempty, where a ladder is a
sequence $E:\N\to\R$ with $E(n)\in S.\mathrm{Energy}$, $E(n)>0$ and $2\cdot E(n+1)\le E(n)$ for all
$n$.
\end{theorem}

\emph{Proof idea.} Forward, the ladder is built by choice, mirroring the zero-extraction of the
Riemann branch; conversely a ladder refutes any candidate $\Delta$ by a geometric majorant. As long
as the ladder lives in $\R$, acyclicity cannot touch it.

The engine is killed exactly when the ladder can be \emph{reflected into the natural numbers}. This
is the physical input of the branch, and it is genuinely open.

\begin{definition}[\code{QuantizationLaw}]\label{def:QuantizationLaw}
\red{}\ A red per-model input (the \code{EnergyBalanceLaw} pattern): every positive state carries a
natural rank, and multiplicative descent of energy is reflected into a \emph{strict} descent of
rank,
\[
  \mathrm{QuantizationLaw}(S):=\exists\,\mathrm{rank}:\mathrm{PositiveState}(S)\to\N,\
    \forall x,y,\ 2\cdot(y:\R)\le(x:\R)\Rightarrow\mathrm{rank}(y)<\mathrm{rank}(x).
\]
\end{definition}

Under the law, the ladder turns from real into natural, and breaks off.

\begin{theorem}[\code{no\_quantizedLadder}]\label{thm:no_quantizedLadder}
\green{}\ A quantized ladder does not exist:
$\mathrm{QuantizationLaw}(S)\wedge\mathrm{GaplessLadder}(S)\Rightarrow\bot$.
\end{theorem}

\emph{Proof idea.} A quantized ladder would be an infinite strictly descending chain of natural
ranks --- Euclid's perpetual engine --- killed by \code{no\_infinite\_descent}
(\cref{sec:engine}). The killer here is the pure well-foundedness of $\N$; no fake ``there are no
engines'' hypothesis is mixed in.

\begin{theorem}[\code{massGap\_of\_quantizationLaw}]\label{thm:massGap_of_quantizationLaw}
\green{}\ Quantization implies the mass gap: $\mathrm{QuantizationLaw}(S)\Rightarrow\mathrm{MassGap}(S)$.
\end{theorem}

\emph{Proof idea.} No gap gives masslessness (\cref{thm:not_massGap_iff_gapless}), masslessness a
ladder (\cref{thm:gapless_iff_nonempty_ladder}), quantization turns it into a natural descent, and
\cref{thm:no_quantizedLadder} closes the contradiction. This is the green conditional theorem of the
branch: spectral quantisation $\Rightarrow$ mass gap.

As with P/NP, one is tempted to add a fourth boundary of the first cause; the machine forbids it,
and the reason runs deeper here.

\begin{theorem}[\code{quantizationLaw\_iff\_massGap}]\label{thm:quantizationLaw_iff_massGap}
\green{}\ For every model the quantization law is equivalent to the presence of a gap, green and
with no boundary anywhere: $\mathrm{QuantizationLaw}(S)\iff\mathrm{MassGap}(S)$.
\end{theorem}

This is the shape of the \emph{condemned} bridge, not of the honest Riemann mirror: there the
equivalence ``law $\iff$ goal'' held only \emph{under} the boundary, here it is green on its own.
Hence to decree the quantization law would be to rename the goal, not to add a cause. The trilemma
confirms this from every side: the universal form is refutable --- the forged massless model
$\{0\}\cup\{2^{-n}\}$ carries an explicit ladder \code{cookedLadder}, and the decree would blow up
the quarantine (\code{ymManifestationLaw\_refutes\_boundary}, \green{}); the existential form is
already proven by the forged gapped model $\{0,1\}$, so the decree would be empty. What is missing is
not a proposition but a \red{} data anchor --- a constructed spectrum or Hamiltonian of a genuine
non-abelian theory, which mathlib does not have.

The branch's only \yellow{} content is what the decree asserts on its own scale. At $A\le 4$ the
supply of deviations is real (\code{smallScale\_deviationSupply}, \green{}); at the decreed
$A\ge 5$ there is no supply at any threshold
(\code{decreedScale\_no\_deviationSupply}, \yellow{}, conditional on \axiom{}): the world of the
first cause is gapped in the language of supplies. An intentional explosion detector,
\code{quarantine\_inconsistent\_if\_supply\_at\_every\_scale} (\yellow{}), keeps the point of
explosion visible.

The epistemic complement (\srcpath{Engine/YangMillsEpistemic.lean}) closes the internal door as for
Collatz and P/NP: self-grounding the quantization law self-destructs
(\code{no\_internalisedYMGround}, \green{}), whence \code{ymCause\_unknowable} (\green{}). What pays
for the contradiction is a genuine construction: from the absence of a gap a halving ladder is built
and carries a real perpetual engine (\code{not\_massGap\_carries\_real\_engine}, \green{}), lawful on
$\R$; the wall arises only when the rank translates it into an $\N$-descent burnt against
\code{no\_perpetual\_engine\_on\_nat}. Everything is gathered in
\code{ym\_locked\_behind\_engine\_status} (\green{}); the repository's taint does not grow.

\begin{remark}
This is \emph{not} a solution of the Clay problem and \emph{not} G\"odel: no incompleteness, no fixed
point, only well-foundedness, and all the epistemics is model-internal. About the genuine quantum
Yang--Mills theory the theorems say nothing; it remains \red{}. The conditionality is named: the
quantization law is a \red{} per-model input, and everything ``under the law'' is an implication;
the tautologization through \cref{thm:quantizationLaw_iff_massGap} is not hidden --- it is exactly
why the only exit is external.
\end{remark}

\subsection{Hodge: quantised charges and payment by cycles}

We read a Hodge class as a \emph{quantised charge} --- a rational cohomology class of aligned type
$(p,p)$ --- and an algebraic cycle as its \emph{payment}. The Hodge conjecture is then completeness:
every quantised aligned charge is paid; an unpaid charge would open an infinite strict descent along
a natural height, a forbidden perpetual engine. Again the honest boundary: there is \emph{no} Hodge
theory in mathlib --- no Hodge classes, no algebraic cycles, no cohomology of varieties. We work
with an abstract ledger \code{HodgeLedger} (classes, predicates \code{IsHodge} and \code{IsAlgebraic},
a height $\mathrm{height}:\mathrm{Cls}\to\N$, and the quantisation anchor ``paid $\iff$ height $0$'').
We do not solve the problem; the remainder is a named input \code{DescentLaw}.

The branch's headline is that here the engine is dead \emph{unconditionally}.

\begin{theorem}[\code{isEmpty\_unpaidDescentChain}]\label{thm:isEmpty_unpaidDescentChain}
\green{}\ For every Hodge ledger $S$ the type $\mathrm{UnpaidDescentChain}(S)$ is empty: an infinite
strictly height-descending chain of unpaid charges exists in no model whatsoever. The axiom list is
just \code{[propext, Quot.sound]}, without even choice --- pure descent at $A=1$.
\end{theorem}

Unlike Yang--Mills, where the real ladder lived at masslessness and was killed by a
\emph{decreeable} quantisation law, here quantisation is built into the model --- the height is a
natural number --- so the engine is dead in every model. What remains contested is not whether the
charge is quantised (yes, by construction) but the existence of the payment \emph{steps}.

\begin{theorem}[\code{hodgeProperty\_of\_descentLaw}]\label{thm:hodgeProperty_of_descentLaw}
\green{}\ If $\mathrm{DescentLaw}(S)$, then $\mathrm{HodgeProperty}(S)$: the descent law implies the
model's Hodge conjecture.
\end{theorem}

\emph{Proof idea.} Strong induction on the height, no choice needed; the same follows by the chain
route through \code{descentSeq}. This is the green conditional theorem of the branch: the descent
law is payment by cycles.

\begin{theorem}[\code{no\_unpaid\_lawful\_model}]\label{thm:no_unpaid_lawful_model}
\green{}\ There is no ledger $S$ satisfying both $\mathrm{DescentLaw}(S)$ and
$\mathrm{Nonempty}(\mathrm{UnpaidClass}(S))$: a model of ``the law plus an unpaid charge'' does not
exist.
\end{theorem}

The converse is honestly vacuous: \code{descentLaw\_of\_hodgeProperty} is proven by an empty
quantifier (if everything is paid, there are no unpaid charges), unlike Yang--Mills where the reverse
rank was built for real. Hence the decisive audit.

\begin{theorem}[\code{descentLaw\_iff\_hodgeProperty}]\label{thm:descentLaw_iff_hodgeProperty}
\green{}\ For every ledger $S$: $\mathrm{DescentLaw}(S)\iff\mathrm{HodgeProperty}(S)$, green and with
no boundary anywhere.
\end{theorem}

Once again the condemned bridge: to decree the law would be to decree the conjecture verbatim. There
can be no sixth field of the decree; what is missing is a \red{} data anchor --- genuine rational
$(p,p)$-classes. The trilemma confirms it, and one face is unique here. Transplanting the
manifestation form from Yang--Mills over \emph{chains} degenerates into a green theorem.

\begin{theorem}[\code{hodgeChainManifestationLaw\_green}]\label{thm:hodgeChainManifestationLaw_green}
\green{}\ $\mathrm{HodgeChainManifestationLaw}$: for every ledger $S$ and every chain
$C:\mathrm{UnpaidDescentChain}(S)$ one has $\mathrm{ChainManifests}(C)$. Since by
\cref{thm:isEmpty_unpaidDescentChain} no such chain exists, the law over chains is vacuously true.
\end{theorem}

This is a vacuity of the ``empty-witness'' type, not an incompatibility: from an empty type of
witnesses one cannot assemble the third trilemma face, which needs a witness \emph{presented} in
green. The honest witness here is not a chain but a single unpaid class, \code{cookedUnpaidClass},
presented by construction without choice.

\begin{theorem}[\code{hodgeManifestationLaw\_refutes\_boundary}]\label{thm:hodgeManifestationLaw_refutes_boundary}
\green{}\ If $\mathrm{HodgeManifestationLaw}$ holds, then
$\neg\,\mathrm{TheStrictLastStep00Obligation}$: the manifestation form of the law is incompatible
with the accepted quarantine boundary.
\end{theorem}

The branch's single \yellow{} is a tripwire,
\code{quarantine\_inconsistent\_if\_hodgeManifestationLaw\_decreed} (conditional on \axiom{}): a
decree of the manifestation form would blow up the quarantine exactly here, and its safety has been
verified.

The epistemic complement (\srcpath{Engine/HodgeEpistemic.lean}) closes the internal door. Self-grounding
the per-model descent law self-destructs (\code{no\_internalisedHodgeGround}).

\begin{theorem}[\code{hodgeCause\_unknowable}]\label{thm:hodgeCause_unknowable}
\green{}\ For every ledger $S$: $\neg\,\mathrm{InternalKnowledgeOfHodgeCause}(S)$ --- internal
self-grounding of the descent law is impossible in every model, unconditionally and without any
axioms.
\end{theorem}

The contradiction has a second, engine route: under the law an unpaid charge unfolds into a genuine
infinite descent chain (\code{unpaidDescentChain\_of\_descentLaw}) and is projected by the height into
a perpetual engine on $\N$ (\code{unpaidClass\_under\_law\_carries\_perpetual\_engine}), built
genuinely rather than \emph{ex falso} (\code{internalisedHodgeGround\_builds\_engine}) and burnt by
\code{no\_perpetual\_engine\_on\_nat}. The universal form is already refuted by forging
(\code{hodgeUniversal\_forged\_refutation}), so the only living option is the per-model input --- which
by \cref{thm:descentLaw_iff_hodgeProperty} equals the goal verbatim.

\begin{theorem}[\code{hodge\_locked\_behind\_engine\_status}]\label{thm:hodge_locked_behind_engine_status}
\green{}\ For every ledger $S$:
\begin{equation}\label{eq:hodge_locked}
\begin{aligned}
  &(\neg\,\mathrm{HodgeDescentLawUniversal})\ \wedge\ \mathrm{IsEmpty}(\mathrm{UnpaidDescentChain}(S))\\
  &\wedge\ (\neg\,\mathrm{InternalKnowledgeOfHodgeCause}(S))\ \wedge\ \bigl(\mathrm{DescentLaw}(S)\iff\mathrm{HodgeProperty}(S)\bigr).
\end{aligned}
\end{equation}
The universal refuted, chains empty in every model, internal knowledge impossible, and the per-model
law equivalent to the model's conjecture --- entirely green; the repository's taint does not grow.
\end{theorem}

\begin{remark}
This is \emph{not} a solution of the Hodge conjecture and \emph{not} G\"odel: no incompleteness, no
fixed point, only model-internal epistemics and $\N$-descent. The self-grounding pair is formally
``ground $+$ its negation''; its substance is paid for by the green identification of both sides
with content (the law $\iff$ the model's conjecture by \cref{eq:hodge_locked}; the negation $\iff$
the presentability of an unpaid charge, \code{not\_hodgeProperty\_iff\_unpaidClass}) and by the
genuine engine route. Hodge's colour on the shared rule: when quantisation is built into the model,
the engine is dead unconditionally and the conjecture shrinks to a single question --- do the payment
steps exist.
\end{remark}
