\section{Coda: the structure of spacetime}\label{sec:coda}

This section adds no essential mathematics: it gathers the meaning of what is already proved. Every
result it leans on was established in \cref{sec:engine,sec:carrier,sec:reduction,sec:firstcause,sec:riemann,sec:pnp,sec:ymhodge,sec:ns,sec:zoo,sec:geometry,sec:bsd,sec:collatz}; the cosmological reading below is a
translation of those results, not an addition, and we keep the line between the rigorous and the
interpreted in plain view.

\paragraph{One prohibition, seven masks.}
Under each of the seven questions --- twins, Riemann, $\mathrm{P}$ versus $\mathrm{NP}$,
Yang--Mills, Navier--Stokes, Hodge, Collatz --- we found the same forbidden object: Euclid's
perpetual engine, an infinite strictly descending chain on a well-ordered line, which does not
exist (\cref{thm:no_infinite_descent}). Each ``deviation'' from the norm is a disguised engine, and
the prohibition kills it: finiteness of the twins, a zero off the critical line, full accounting by
fast motion, a gapless spectrum, a singular cascade, a permitted charge without a carrier, a
Collatz orbit that never falls into the vacuum. Seven masks, one face. Everywhere the same
watershed of honesty: the \emph{structural} half --- ``where the engine prohibition holds, there is
no deviation'' --- is \green{}~machine-proved; the \emph{attachment} to the genuine analytic object
is either \yellow{}~conditional on the single first cause \axiom{} or left as an \red{}~open input.
Not a single millennium problem is solved. Something else is proved, and it may matter more: they
are one problem.

An eighth, Birch--Swinnerton-Dyer (\cref{sec:bsd}), enters differently --- there the engine is not a
guard but the \emph{method}: Fermat's descent itself bounds the rational points of an elliptic
curve, and the rank parity $(-1)^{\mathrm{rank}}$ is the same invariant that stands behind Riemann,
while the analytic bridge stays honestly open. Beyond the masks lie the open neighbours (Chowla,
abc, Beal, Lehmer, Landau) and the arithmetic zoo (Polignac's cousins, Sophie Germain, Goldbach,
Legendre, perfect and Fermat numbers): the same device, a real proved anchor read through the
engine plus a named red gate. One pearl fell out \green{}~and unconditional along the way ---
Sophie Germain primes with $p \equiv 3 \pmod 4$ divide the corresponding Mersenne numbers, an
Euler--Lagrange classic, machine-checked.

\paragraph{Space, the arrow of time, the singularity.}
The unity goes deeper than a coincidence of form. The number line with its strict order of
traversal is \emph{space}: the set of states, with a singularity-bottom at zero. The engine's
irreversibility is \emph{time}: \code{engine\_never\_returns} proves height is \code{StrictAnti}, so
there is no way back --- the arrow of time is a theorem, not a postulate. It has a beginning (the
event $0 \to 1$, which cannot be caused from inside, since self-ignition would be a perpetual
engine, \code{no\_internalisedOriginEvent}) and an unreachable end (infinity, which only the
forbidden engine could reach); and there is no tower of universes beneath the first cause, for the
regress of causes is well-founded (\code{no\_rankedMetaFractalBranch}). Hence the main theorem in
its sharpest form: one cannot know that the twins are infinite; but if the unknowable first cause is
accepted as truth, they are infinite, and rigorously so. Inside a universe defined by the strict
order of traversal the first cause can be neither derived nor known --- both cost a perpetual engine
--- it can only be accepted from outside.

The shape of this structure is an \emph{inverted} fractal. A classical fractal runs outward, gaining
detail toward larger scales; the fractal of Euclid's path is contractive --- the same descent
pattern repeats at ever finer scales as every trajectory coils toward the origin. Not one trajectory
escapes; all coil back. The impossibility of the engine is precisely ``no escape,'' and the
infinitude of the twins is the one thing that cannot be certified from inside the contracting order.

\paragraph{Why this resembles a theory of everything.}
Here we look into the distance, clearly marking it as \emph{interpretation}, not theorem. From the
single prohibition --- nothing from nothing, no descending forever --- a surprising amount of
\emph{structure} follows at once: the arrow of time (\cref{sec:engine}); a beginning, the
singularity from which it is launched (\cref{sec:firstcause}); quantisation, hence mass, since a
gapless ladder would be an engine (\cref{sec:ymhodge}); stability of the vacuum, with no second
bottom; a finite price of computation, Landauer's principle in rank clothing (\cref{sec:pnp});
completeness of the charge spectrum (\cref{sec:ymhodge}); smoothness of flow (\cref{sec:ns}); and
realness of the spectrum (\cref{sec:firstcause}). Nearly everything we call the structure of the
world turns out to be one conservation law read on different objects --- the very gesture of a
theory of everything. This reading resonates with causal set theory (spacetime as a locally finite
poset, order as causal time) and with ``it from bit'' and Landauer's principle (information is
physical, and carries an energy price); the engine prohibition is the first law of thermodynamics
lifted from machines to the very form of being.

Two further correspondences suggest themselves in the same interpretive spirit. The combinatorial
curvature of the descent graph (\cref{sec:geometry}) is read off the branching, not off any body
sitting in it, so it contributes to the geometry with no visible carrier of mass --- the role of
\emph{dark matter}, itself inferred only from its contribution to curvature. And an observer at the
centre of his horizon, into whom the arrow of time runs from all sides, sees the far centres recede
as though the recession accelerated; yet nothing is pushed apart --- the apparent acceleration is an
artifact of the curvature of the contracting order, not a dark energy, echoing the timescape and
backreaction cosmologies~\cite{wiltshire2007} in which acceleration follows from inhomogeneity and
curvature. Neither is offered as physics.

We are emphatic about the boundary. We have solved no millennium problem and declare none solved:
\code{twin\_prime\_conjecture} remains a \code{sorry}, and the hard halves of all branches are
either \yellow{}~conditional on \axiom{} or \red{}~open. This is not a physical theory of everything
--- it predicts no constant, gives no equations of motion, is not testable. The ``theory of
everything'' here is a structural observation: one impossibility organises all seven problems, and
that suffices for the shape of spacetime to be read off from it. This is expressly \emph{not} a
G\"odelian incompleteness phenomenon~\cite{godel1931}: the obstruction is not the undecidability of
a formal system but a conservation law --- the first cause is unknowable from inside because knowing
it would run the forbidden engine, not because the theory fails to decide it.

\paragraph{The space of all spaces.}
One firm rigorous skeleton underlies the long view. The engine is defined not on the twins nor on
$\N$ but over \emph{any} relation, and its impossibility is intrinsic: a well-founded relation carries
no perpetual engine (\cref{thm:no_perpetual_engine_of_wellFounded}), and the prohibition transfers
along any rank into a well-founded order (\cref{thm:no_perpetual_engine_of_rank}) --- in particular
along a height into $\N$. From here, all seven masks are \emph{one} theorem. The heights of the twins, the Liouville rank, the
quantised Hodge charge, the Yang--Mills ladder, the shells of the dyadic cascade are all
$\N$-valued, and the engine prohibition transfers to each by the same transport of
\cref{thm:no_perpetual_engine_of_rank}. We did not prove seven times; we proved once and
transferred.

And then something surfaces worth saying aloud. Every space on which analysis stands ---
$\R$, $\R^n$, the complex plane, metric and function spaces --- is built out of the number line,
$\N \to \Z \to \mathbb{Q} \to \R$, so the well-ordered line and with it the engine prohibition lie
at their very foundation. In this sense Euclid's path is not one of the spaces but the space of all
spaces. Yet the order of the continuum $\R$ is \emph{not} well-founded: on it the engine
\emph{works} (\code{perpetualEngine\_on\_real}, the chain $(1/2)^n$ that screws toward zero without
reaching it). This is not an escape but the reverse side: \code{universal\_engine\_dividing\_line}
says exactly where motion is forbidden (the discrete, well-founded) and where it is realised (the
continuum) --- and where realised, we already know what it is: mass, singularity, black hole. To
remove the prohibition is to remove well-foundedness, hence $\N$, hence numbers and mathematics
itself; the prohibition is not an added axiom but the form of counting.

\paragraph{Finale.}
Euclid's perpetual engine is, in essence, Fermat's infinite descent, and that goes back to the
\emph{Elements}~\cite{euclid-elements}: the oldest demonstrative object of mathematics, the idea that
natural numbers cannot be decreased forever. Twenty-three centuries later the same object turned out
to be the skeleton of physics' youngest question. Pythagoras said all is number; Euclid's path
refines the formula --- all is a well-ordered march that cannot be turned back and cannot be walked
to the end, and that is time, while that along which it goes is space. The engine is impossible, and
its impossibility is not a fact about machines but a form of existence. Beyond that is the silence of
the proven: \code{twin\_prime\_conjecture} awaits its line, the tripwires against a refuted boundary
are armed, and the single axiom bears exactly as much as we have honestly laid upon it. The rest is
open, and that too is part of the honesty.
