\documentclass[11pt,a4paper]{article}
\input{macros}
\usepackage[english]{babel}

\title{The Perpetual Engine of Euclid:\\
  A Machine-Verified Reading of the Great Conjectures\\
  through a Single Axiom}
\author{D. S. Semenov \and L. R. Galieva}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
We formalize, in the Lean~4 proof assistant, a programme organised around a single
physical prohibition: the impossibility of a \emph{perpetual engine} --- Fermat's infinite
descent, restated for the centres of the Euclidean pairs $6m\pm1$. From this one fact we
obtain a machine-checked reduction of the twin-prime conjecture to a \emph{single} open node,
and we read a family of Millennium-scale questions --- the Riemann hypothesis, $\mathrm{P}$
versus $\mathrm{NP}$, Yang--Mills, Hodge, Navier--Stokes, Birch--Swinnerton-Dyer, together
with Mersenne primes, the Collatz problem, and an arithmetic zoo --- as shadows of the same
prohibition on different objects. The node cannot be closed from within, since its
self-grounding would build the forbidden engine; it is therefore accepted \emph{from outside}
by a single quarantined axiom, the \emph{first cause}. Our main theorem
(``higher-energy incompatibility'') shows that the infinitude of the twins can be neither
proved nor refuted inside the system without such an engine. We are scrupulous about what is
and is not proved: the whole development rests on \emph{exactly one} non-standard axiom and
contains \emph{exactly one} \code{sorry} (the twin-prime conjecture itself); a machine verifier
marks every axiom-tainted declaration ($47$ at present) and re-checks these two invariants on
every run. \textbf{No Millennium problem is solved or claimed to be solved.} Every formal
statement in this article carries the name of the corresponding machine-verified declaration;
the complete Lean source is public, and the identical content is browsable online.
\end{abstract}

\begin{center}\small
\fbox{\begin{minipage}{0.92\linewidth}\centering
Machine-verified source (Lean~4): \url{https://github.com/elamaunt/Euclids-path}\\
The same content, as a bilingual (English/Russian) site:
\url{https://elamaunt.github.io/Euclids-path/}\\[2pt]
Every theorem below names its Lean declaration; run \code{lake env lean scripts/VerifyAll.lean}
to re-check the honesty invariants.
\end{minipage}}
\end{center}

\medskip
\noindent\textbf{Keywords.} twin primes; infinite descent; formalization; Lean~4; first cause;
Riemann hypothesis; $\mathrm{P}$ versus $\mathrm{NP}$; Navier--Stokes; Collatz.

\smallskip
\noindent\textbf{MSC 2020.} 11A41, 11N05, 68V20, 03F03, 11P32.

\tableofcontents

\input{sections/01-introduction.en}
\input{sections/02-engine.en}
\input{sections/03-carrier.en}
\input{sections/04-reduction.en}
\input{sections/05-firstcause.en}
\input{sections/06-riemann.en}
\input{sections/07-pnp.en}
\input{sections/08-yangmills-hodge.en}
\input{sections/09-navierstokes.en}
\input{sections/10-zoo.en}
\input{sections/11-geometry.en}
\input{sections/12-bsd-neighbours.en}
\input{sections/13-collatz.en}
\input{sections/14-coda.en}
\input{sections/15-status.en}
\input{sections/16-figures.en}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}
